---
title: "How to make a reasonable person"
date: "May 8, 2026"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    toc_float: true
    theme: cosmo
    highlight: tango
---

```{r setup, include = F, message = F, warning = F}
require(knitr)         # for knitting  
library(ggstatsplot)   # extension of ggplot2
library(lme4)          # for mixed effects models 
library(lmerTest)
library(DT)            # for datatable() function 
library(brms)          # for Bayesian regression models
library(emmeans)       # for estimated marginal means
library(Metrics)       # for RMSE
library(Hmisc)         # for bootstrapped confidence intervals
library(zeallot)       # for unpacking lists
library(scales)        # hue palette
library(showtext)      # for fonts
library(broom.mixed)   # for augmenting fitted models
library(ggtext)        # for text formatting in ggplot
library(ggrepel)       # avoid overlap in geom_text
library(patchwork)     # for multiple plots
library(ggh4x)         # for facet_wrap2 with axes on all panels
library(tidyverse)     # for everything else


knitr::opts_chunk$set(echo = T, warning = F, message = F)

# set defaults
theme_set(theme_classic() + 
            theme(text = element_text(size = 14))) 
options(dplyr.summarise.inform = F)
options(xtable.floating = F)
options(xtable.timestamp = "")
options(readr.show_col_types = F)

data_dir = '../data/'
figures_dir = '../figures/'

set.seed(1)
```

# Helper functions

```{r}
feature_types = c('generic', 'demographics', 'dispositions', 'actions')
feature_type_colors = c("generic" = "gray50",
                        "demographics" = "#88CCEE",
                        "dispositions" = "#EE8866",
                        "actions" = "#DDCC77")
features = read_csv(str_c(data_dir, 'features_mapping.csv')) %>%
  mutate(feature_key = factor(feature_key),
         feature_type = factor(feature_type, levels = feature_types))

conditions = c("relevance", "controllability", "normality",
               "undo", "same", "different", "endorse")

condition_labels = c(
  "relevance" = "relevance",
  "controllability" = "controllability",
  "normality" = "normality",
  "undo" = "What could be changed <br> &emsp;&emsp;&ensp;to <b>undo</b> the outcome?",
  "same" = "What would be the <b>same</b> <br> in a reasonable person?",
  "different" = "What would be <b>different</b> <br> in a reasonable person?",
  "endorse" = "What subjective standard <br> would you <b>endorse</b>?"
)

condition_titles = c(
  "relevance" = "<b>relevance</b>",
  "controllability" = "<b>controllability</b>",
  "normality" = "<b>normality</b>",
  "undo" = "<b>undo</b> the<br>outcome",
  "same" = "<b>same</b> as<br>reasonable",
  "different" = "<b>different</b> from<br>reasonable",
  "endorse" = "<b>endorsed</b><br>standard"
)

attention.key = read_csv(str_c(data_dir, 'attention_key.csv'))
```

## Rounding
```{r include = F}
# round percentages ensuring they sum to 100% using largest remainder method
round_percent = function(x) {
  pct = x * 100
  floored = floor(pct)
  remainder = pct - floored
  shortfall = 100 - sum(floored)
  idx = order(remainder, decreasing = TRUE)[seq_len(shortfall)]
  floored[idx] = floored[idx] + 1
  paste0(floored, "%")
}
```

## Demographics
```{r include = F}
print_demographics = function(data) {
  # gender
  data %>%
    group_by(gender) %>%
    summarise(n = n()) %>%
    print()
  
  # age
  print('age:')
  mean(data$age, na.rm = T) %>% print()
  sd(data$age, na.rm = T) %>% print()
  
  # race
  data %>%
    group_by(race) %>%
    summarise(n = n()) %>%
    print()

  # time taken
  print('time taken (min):')
  print(mean(data$time, na.rm = T)/60)
  print(sd(data$time, na.rm = T)/60)
}
```

## Process data
```{r}
get_tense = function(condition) {
  feature_tense = case_when(
    condition == 'controllability' ~ 'gerund',
    condition %in% c('normality', 'relevance', 'undo') ~ 'present',
    condition %in% c('same', 'different') ~ 'past',
    condition == 'endorse' ~ 'person'
    )
  return(feature_tense)
}

process_data = function(condition, data_dir, shift = 0) {
  # import raw data
  data = read_csv(str_c(data_dir, condition, '/trials.csv'))
  data.participants = read_csv(str_c(data_dir, condition, '/participants.csv'))
  print(str_c('Recruited n = ', length(unique(data$id))))

  # filter attention checks
  excluded_ids = data %>%
    distinct(id, story, attention) %>%
    left_join(attention.key, by = 'story') %>%
    mutate(attention_correct = (attention == key)) %>%
    group_by(id) %>%
    summarise(correct = sum(attention_correct)) %>%
    filter(correct < 9) %>%
    .$id
  
  # process trial data
  data = data %>%
    filter(!(id %in% excluded_ids)) %>%
    select(-attention, -story_order, -feature_order) %>%
    left_join(features %>%
                select(feature_key,
                       feature = !!str_c('feature_', get_tense(condition)),
                       feature_type),
              by = 'feature') %>%
    select(-feature) %>%
    mutate(feature_type = factor(feature_type,
                                 levels = feature_types,
                                 ordered = T),
           id = as.integer(factor(id)) + shift,
           condition = condition)
  print(str_c('Included n = ', length(unique(data$id))))
  
  # process participant data
  data.participants = data.participants %>%
    filter(!(id %in% excluded_ids)) %>%
    mutate(id = as.integer(factor(id)) + shift)

  return(list(data, data.participants))
}
```


# Experiment 1

## Data

### Relevance
```{r}
c(data.rel, data.rel.participants) %<-%
  process_data('relevance', str_c(data_dir, 'experiment1/'))

data.rel.participants %>%
  select(id, feedback) %>%
  datatable(rownames = F)
```

### Controllability
```{r}
c(data.ctrl, data.ctrl.participants) %<-%
  process_data('controllability', str_c(data_dir, 'experiment1/'), 50)

data.ctrl.participants %>%
  select(id, feedback) %>%
  datatable(rownames = F)
```

### Normality
```{r}
c(data.norm, data.norm.participants) %<-%
  process_data('normality', str_c(data_dir, 'experiment1/'), 100)

data.norm.participants %>%
  select(id, feedback) %>%
  datatable(rownames = F)
```

## Demographics
```{r}
print_demographics(data.ctrl.participants %>%
                     rbind(data.norm.participants) %>%
                     rbind(data.rel.participants))
```

## Analysis

### Marginal means

```{r}
# combine predictor data with factors
predictor_data = data.rel %>%
  rbind(data.ctrl) %>%
  rbind(data.norm) %>%
  mutate(condition = factor(condition, levels = conditions),
         id = factor(id))

# mixed effects model
predictor_model = lmer(response ~ condition * feature_type + (1 + feature_type | id),
                       data = predictor_data)

# estimate marginal means with CIs
predictor_emm = emmeans(predictor_model, ~ condition | feature_type)
predictor_summary_emm = as.data.frame(predictor_emm) %>%
  rename(y = emmean, ymin = asymp.LCL, ymax = asymp.UCL) %>%
  select(condition, feature_type, y, ymin, ymax)

predictor_summary_emm %>%
  mutate(
    condition = factor(condition, levels = conditions),
    CI = sprintf("[%.2f, %.2f]", ymin, ymax),
    mean = round(y, 2)
  ) %>%
  arrange(condition, feature_type) %>%
  select(condition, feature_type, mean, CI)
```

### Pairwise differences

```{r}
predictor_emm_by_condition = emmeans(predictor_model, ~ feature_type | condition)
predictor_contrasts = pairs(predictor_emm_by_condition, adjust = "bonferroni")
summary(predictor_contrasts)
```

## Figures

```{r fig.width = 5, fig.height = 3}
ggplot(data = predictor_summary_emm,
       aes(x = feature_type,
           y = y,
           color = feature_type)) +
  geom_linerange(aes(ymin = ymin, ymax = ymax,
                     color = feature_type),
                 linewidth = 0.8,
                 show.legend = F) +
  geom_point(size = 3) +
  facet_wrap(~ condition, scales = 'free_y',
             labeller = as_labeller(condition_labels),
             strip.position = "bottom") +
  scale_y_continuous(name = "rating",
                     limits = c(1, 7),
                     breaks = 1:7,
                     labels = c("1", "", "", "4", "", "", "7")) +
  scale_color_manual(values = feature_type_colors) +
  theme(text = element_text(size = 12),
        legend.position = 'inside',
        legend.position.inside = c(0.5, -0.18),
        legend.direction = "horizontal",
        legend.title = element_blank(),
        legend.text = element_text(
          size = 12, margin = margin(0)),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 12),
        strip.background = element_blank(),
        strip.placement = "outside",
        strip.text = element_text(
          size = 12, face = 'bold',
          margin = margin(t = 0, b = 0.1, unit = "cm")),
        panel.grid.major.y = element_line(),
        plot.margin = margin(t = 0.1, b = 0.8, unit = "cm"))

# ggsave(str_c(figures_dir, 'exp1.pdf'),
#        width = 5, height = 3)
```


# Experiment 2

## Data

### Undo
```{r}
c(data.undo, data.undo.participants) %<-%
  process_data('undo', str_c(data_dir, 'experiment2/'))

data.undo.participants %>%
  select(id, feedback) %>%
  datatable(rownames = F)
```

### Same
```{r}
c(data.same, data.same.participants) %<-%
  process_data('same', str_c(data_dir, 'experiment2/'), 50)

data.same.participants %>%
  select(id, feedback) %>%
  datatable(rownames = F)
```

### Different
```{r}
c(data.diff, data.diff.participants) %<-%
  process_data('different', str_c(data_dir, 'experiment2/'), 100)

data.diff.participants %>%
  select(id, feedback) %>%
  datatable(rownames = F)
```

### Endorse
```{r}
c(data.endorse, data.endorse.participants) %<-%
  process_data('endorse', str_c(data_dir, 'experiment2/'), 150)

data.endorse.participants %>%
  select(id, feedback) %>%
  datatable(rownames = F)
```

## Demographics
```{r}
print_demographics(data.undo.participants %>%
                     rbind(data.same.participants) %>%
                     rbind(data.diff.participants) %>%
                     rbind(data.endorse.participants))
```

## Analysis

### Descriptive statistics

Selection rates (undo, same, different):
```{r}
data.selection_total = bind_rows(data.undo, data.same, data.diff) %>%
  mutate(condition = factor(condition, levels = conditions)) %>%
  group_by(condition) %>%
  summarise(prop_selected = mean(response))

data.selection_total
```

Selection distribution by feature type:
```{r}
data.selection_prop = bind_rows(data.undo, data.same, data.diff) %>%
  mutate(condition = factor(condition, levels = conditions)) %>%
  group_by(condition, feature_type) %>%
  summarise(n_selected = sum(response),
            n = n(),
            .groups = 'drop') %>%
  group_by(condition) %>%
  mutate(prop = n_selected / sum(n_selected))
data.selection_prop
```

Pairwise correlations:
```{r}
selection_rates = bind_rows(data.undo, data.same, data.diff) %>%
  group_by(condition, story, feature_key) %>%
  summarise(selection_rate = mean(response),
            .groups = 'drop') %>%
  pivot_wider(names_from = condition,
              values_from = selection_rate)

# undo vs. same
round(cor(selection_rates$undo, selection_rates$same, use = "complete.obs"), 3)

# undo vs. different
round(cor(selection_rates$undo, selection_rates$different, use = "complete.obs"), 3)

# same vs. different
round(cor(selection_rates$same, selection_rates$different, use = "complete.obs"), 3)
```

Agreement ratings (endorse):
```{r}
data.endorse %>%
  group_by(feature_type) %>%
  summarise(
    mean = mean(response),
    sd = sd(response),
    ci_low = mean_cl_boot(response)$ymin,
    ci_high = mean_cl_boot(response)$ymax,
    n = n()
  )
```

### Models

```{r}
data.modeling = bind_rows(data.undo, data.same, data.diff, data.endorse) %>%
  left_join(data.ctrl %>%
              group_by(story, feature_key) %>%
              summarise(controllability = mean(response)) %>%
              left_join(data.norm %>%
                          group_by(story, feature_key) %>%
                          summarise(normality = mean(response)),
                        by = c('story', 'feature_key')) %>%
              left_join(data.rel %>%
                          group_by(story, feature_key) %>%
                          summarise(relevance = mean(response)),
                        by = c('story', 'feature_key')),
            by = c('story', 'feature_key')) %>%
  mutate(story = factor(story),
         id = factor(id))
```

#### Fitting models

```{r}
fit.undo = brm(
  formula = response ~ 1 + relevance + controllability + normality +
    (1 + relevance + controllability + normality | id) + (1 | story),
  data = data.modeling %>% filter(condition == 'undo'),
  family = bernoulli(link = "logit"),
  iter = 2000,
  seed = 1,
  file = "cache/fit.undo"
)

fit.same = brm(
  formula = response ~ 1 + relevance + controllability + normality +
    (1 + relevance + controllability + normality | id) + (1 | story),
  data = data.modeling %>% filter(condition == 'same'),
  family = bernoulli(link = "logit"),
  iter = 2000,
  seed = 1,
  file = "cache/fit.same"
)

fit.diff = brm(
  formula = response ~ 1 + relevance + controllability + normality +
    (1 + relevance + controllability + normality | id) + (1 | story),
  data = data.modeling %>% filter(condition == 'different'),
  family = bernoulli(link = "logit"),
  iter = 2000,
  seed = 1,
  file = "cache/fit.diff"
)

fit.endorse = brm(
  formula = response ~ 1 + relevance + controllability + normality +
    (1 + relevance + controllability + normality | id) + (1 | story),
  data = data.modeling %>% filter(condition == 'endorse') %>% drop_na(),
  iter = 2000,
  seed = 1,
  file = "cache/fit.endorse"
)
```

#### Model coefficients

```{r}
# extract fixed effects and convert to odds ratios
extract_or = function(fit, condition) {
  summary(fit)$fixed[-1, c("Estimate", "l-95% CI", "u-95% CI")] %>%
    rownames_to_column("predictor") %>%
    mutate(across(c(Estimate, `l-95% CI`, `u-95% CI`), exp),
           condition = condition)
}
undo_CI_95 = extract_or(fit.undo, "undo")
same_CI_95 = extract_or(fit.same, "same")
diff_CI_95 = extract_or(fit.diff, "different")

endorse_CI_95 = summary(fit.endorse)$fixed[-1, c("Estimate", "l-95% CI", "u-95% CI")] %>%
  rownames_to_column("predictor") %>%
  mutate(condition = "endorse")

CI_95_all = undo_CI_95 %>%
  rbind(same_CI_95) %>%
  rbind(diff_CI_95) %>%
  rbind(endorse_CI_95) %>%
  rename(mean = Estimate, l95 = `l-95% CI`, u95 = `u-95% CI`) %>%
  mutate(across(c(mean, l95, u95), ~ round(.x, 2)),
         condition = factor(condition, levels = conditions),
         predictor = factor(predictor, levels = conditions))

CI_95_all
```

#### Model predictions
```{r}
get_binary_pred_actual = function(fit, condition_name) {
  pred = fitted(fit, summary = TRUE)[, "Estimate"]
  data = data.modeling %>%
    filter(condition == condition_name) %>%
    mutate(predicted = pred) %>%
    group_by(story, feature_key) %>%
    summarise(predicted = mean(predicted),
              actual = mean(response),
              .groups = 'drop') %>%
    mutate(condition = condition_name)
  return(data)
}

get_continuous_pred_actual = function(fit, condition_name) {
  pred = fitted(fit, summary = TRUE)[, "Estimate"]
  data = data.modeling %>%
    filter(condition == condition_name) %>%
    drop_na() %>%
    mutate(predicted = pred) %>%
    group_by(story, feature_key) %>%
    summarise(predicted = mean(predicted),
              actual = mean(response),
              .groups = 'drop') %>%
    mutate(condition = condition_name)
  return(data)
}

model_predictions = bind_rows(
  get_binary_pred_actual(fit.undo, "undo"),
  get_binary_pred_actual(fit.same, "same"),
  get_binary_pred_actual(fit.diff, "different"),
  get_continuous_pred_actual(fit.endorse, "endorse")
) %>%
  mutate(condition = factor(condition, levels = conditions))

# Calculate R-squared for each condition
model_metrics = model_predictions %>%
  group_by(condition) %>%
  summarise(r = cor(predicted, actual),
            rmse = sqrt(mean((predicted - actual)^2)),
            .groups = 'drop') %>%
  mutate(label = sprintf("RMSE = %.2f\nR = %.2f", rmse, r))
```


## Figures

### Example (Nadia)

Aesthetics:
```{r fig.width = 11, fig.height = 5}
nadia_ft_map = data.undo %>%
  rbind(data.endorse) %>%
  filter(story == 'Nadia') %>%
  distinct(feature_key, feature_type) %>%
  mutate(color = feature_type_colors[as.numeric(feature_type)])

nadia_ft_colors = setNames(nadia_ft_map$color, nadia_ft_map$feature_key)

nadia_labels_plain = c(
  'generic' = '(generic)',
  'female' = 'female',
  'white' = 'White',
  '34 years old' = '34 years old',
  'red hair' = 'red hair',
  'software engineer' = 'software engineer',
  'timid person' = 'timid',
  'organized person' = 'organized',
  'enjoy oil painting' = 'enjoys oil painting',
  'collect vinyl records' = 'collects vinyl records',
  'attend couples therapy years ago' = 'attended couples therapy',
  'go to emergency room for bruises' = 'went to ER for bruises prior',
  'not seek help after emergency room incident' = 'did not seek help after ER incident',
  'scared when franklin screamed' = 'got scared when Franklin screamed',
  'grab revolver' = 'grabbed revolver',
  'shoot franklin with revolver' = 'shot Franklin with revolver')

# colored labels using element_markdown
nadia_labels = setNames(
  paste0('<span style="color:',
         nadia_ft_colors[names(nadia_labels_plain)],
         '">**', nadia_labels_plain, '**</span>'),
  names(nadia_labels_plain))

nadia_levels = rev(names(nadia_labels))
```

Helper functions:
```{r}
# shared theme for individual condition panels
nadia_theme = theme(
  text = element_text(size = 12),
  plot.title = element_markdown(size = 12, hjust = 0.5),
  legend.position = "none",
  axis.title.y = element_blank(),
  axis.text.y = element_blank(),
  axis.ticks.y = element_blank(),
  axis.text.x = element_text(size = 12),
  panel.grid.major = element_line(linewidth = 0.3, color = "grey90"),
  plot.margin = margin(0))

# for relevance, controllability, and normality
nadia_pred_panel = function(data, cond) {
  d = data %>%
    filter(story == 'Nadia', condition == cond) %>%
    group_by(feature_key) %>%
    summarise(mean = mean(response),
              ci_low = mean_cl_boot(response)$ymin,
              ci_high = mean_cl_boot(response)$ymax,
              .groups = 'drop') %>%
    mutate(feature_key = factor(feature_key, levels = nadia_levels))

  ggplot(d, aes(x = mean, y = feature_key,
                color = feature_key)) +
    geom_pointrange(aes(xmin = ci_low, xmax = ci_high),
                    size = 0.3, linewidth = 0.7) +
    scale_x_continuous(name = "rating", limits = c(0.5, 7.5),
                       breaks = c(1, 4, 7)) +
    scale_y_discrete(limits = nadia_levels,
                     labels = nadia_labels,
                     drop = FALSE) +
    scale_color_manual(values = nadia_ft_colors) +
    ggtitle(condition_titles[cond]) +
    nadia_theme
}

# for undo, same, and different
nadia_sel_panel = function(data, cond) {
  d = data %>%
    filter(story == 'Nadia', condition == cond) %>%
    group_by(feature_key) %>%
    summarise(mean = mean(response),
              ci_low = mean_cl_boot(response)$ymin,
              ci_high = mean_cl_boot(response)$ymax,
              .groups = 'drop') %>%
    mutate(feature_key = factor(feature_key, levels = nadia_levels))

  ggplot(d, aes(x = mean,
                y = feature_key,
                fill = feature_key,
                color = feature_key)) +
    geom_col(alpha = 0.5, width = 0.7) +
    geom_linerange(aes(xmin = ci_low, xmax = ci_high),
                   linewidth = 0.7) +
    scale_x_continuous(name = "selection",
                       limits = c(-0.1, 1.1),
                       breaks = c(0, 0.5, 1),
                       labels = c(0, 0.5, 1)) +
    scale_y_discrete(limits = nadia_levels,
                     labels = nadia_labels,
                     drop = FALSE) +
    scale_fill_manual(values = nadia_ft_colors) +
    scale_color_manual(values = nadia_ft_colors) +
    ggtitle(condition_titles[cond]) +
    nadia_theme
}
```

```{r fig.width = 11, fig.height = 5}
legend_data = tibble(
  feature_type = factor(feature_types[2:4], levels = feature_types[2:4]),
  x = 1, y = 3:1
)

p_legend = ggplot(legend_data,
                  aes(x = x, y = y,
                      color = feature_type)) +
  geom_point(size = 2) +
  geom_text(aes(label = feature_type),
            hjust = 0,
            nudge_x = 0.15,
            size = 4) +
  scale_color_manual(values = feature_type_colors) +
  scale_x_continuous(limits = c(0.9, 2.5)) +
  scale_y_continuous(limits = c(0.5, 3.5)) +
  theme_void() +
  theme(legend.position = "none")

p1 = nadia_pred_panel(data.rel, "relevance") +
  theme(axis.text.y = element_markdown(size = 12)) +
  inset_element(p_legend, left = -2.7, bottom = 0.85, right = -1, top = 1)
p2 = nadia_pred_panel(data.ctrl, "controllability")
p3 = nadia_pred_panel(data.norm, "normality")

p4 = nadia_sel_panel(data.undo, "undo")
p5 = nadia_sel_panel(data.same, "same")
p6 = nadia_sel_panel(data.diff, "different")

p7 = ggplot(data.endorse %>%
              filter(story == 'Nadia') %>%
              group_by(feature_key) %>%
              summarise(mean = mean(response),
                        ci_low = mean_cl_boot(response)$ymin,
                        ci_high = mean_cl_boot(response)$ymax,
                        .groups = 'drop') %>%
              mutate(feature_key = factor(feature_key, levels = nadia_levels)),
       aes(x = mean, y = feature_key,
           color = feature_key)) +
  geom_pointrange(aes(xmin = ci_low, xmax = ci_high),
                  size = 0.3, linewidth = 0.7) +
  scale_x_continuous(name = "agreement",
                     limits = c(-10, 110),
                     breaks = c(0, 50, 100)) +
  scale_y_discrete(limits = nadia_levels,
                   labels = nadia_labels,
                   drop = FALSE) +
  scale_color_manual(values = nadia_ft_colors) +
  ggtitle(condition_titles["endorse"]) +
  nadia_theme

p1 + p2 + p3 + p4 + p5 + p6 + p7 +
  plot_layout(nrow = 1, widths = rep(1, 7))

# ggsave(str_c(figures_dir, "example_nadia.pdf"),
#        width = 11, height = 5)
```

### Aggregate results

#### Selection barplots

For undo, same, and different conditions only.
```{r}
all_colors = c(feature_type_colors,
               'selected' = 'gray75',
               'not selected' = 'gray95')

bar_height = 0.3  # half-height of each bar
```

Top bars: selected (dark gray) vs not selected (light gray)
```{r}
data.top_bar = data.selection_total %>%
  rowwise() %>%
  reframe(condition = condition,
          category = c("selected", "not selected"),
          prop = c(prop_selected, 1 - prop_selected)) %>%
  mutate(condition = factor(condition, levels = conditions),
         ypos = 2) %>%
  arrange(condition, desc(category == "selected")) %>%
  group_by(condition) %>%
  mutate(xmax = cumsum(prop),
         xmin = xmax - prop,
         xmid = (xmin + xmax) / 2) %>%
  ungroup()

top_bar_labels = tibble(
  condition = factor(c("undo", "undo",
                       "same", "same",
                       "different", "different"),
                     levels = conditions),
  label = c("CHANGED", "NOT CHANGED",
            "SAME", "NOT THE SAME",
            "DIFFERENT", "NOT DIFFERENT"),
  x = c(0, 1, 0, 1, 0, 1),
  hjust = c(0, 1, 0, 1, 0, 1),
  label_color = rep(c("gray50", "gray75"), 3)
)
```

Bottom bars: feature type breakdown
```{r}
data.bottom_bar = data.selection_prop %>%
  mutate(category = as.character(feature_type),
         ypos = 1) %>%
  arrange(condition, feature_type) %>%
  group_by(condition) %>%
  mutate(xmax = cumsum(prop),
         xmin = xmax - prop,
         xmid = (xmin + xmax) / 2,
         pct_label = round_percent(prop)) %>%
  ungroup()

bottom_bar_labels = data.selection_prop %>%
  filter(condition == "same") %>%
  ungroup() %>%
  arrange(feature_type) %>%
  mutate(xlabel = c(0.1, 0.63, 1.05),
         label_color = feature_type_colors[as.character(feature_type)])
```

Zoom lines:
```{r}
# left edge (0, top) to (0, bottom)
# right edge (selected, top) to (1, bottom)
data.zoom = data.selection_total %>%
  transmute(condition,
            x_right = prop_selected,
            x_left = 0)
```

Generating plot:
```{r}
p_selections = ggplot() +
  # top bars
  geom_rect(data = data.top_bar,
            aes(xmin = xmin,
                xmax = xmax,
                ymin = ypos - bar_height,
                ymax = ypos + bar_height,
                fill = category),
            color = "gray70", 
            linewidth = 0.4) +
  geom_text_repel(data = data.top_bar,
                  aes(x = xmid,
                      y = ypos,
                      label = percent(prop, accuracy = 1)),
                  direction = "x",
                  point.size = NA,
                  box.padding = 0.1,
                  size = 3.5,
                  color = "black") +
  # bottom bars
  geom_rect(data = data.bottom_bar,
            aes(xmin = xmin,
                xmax = xmax,
                ymin = ypos - bar_height,
                ymax = ypos + bar_height,
                fill = category),
            color = "gray70",
            linewidth = 0.4) +
  geom_text_repel(data = data.bottom_bar,
                  aes(x = xmid,
                      y = ypos,
                      label = pct_label),
                  direction = "x",
                  point.size = NA,
                  box.padding = 0.1,
                  size = 3.5) +
  # zoom lines
  geom_segment(data = data.zoom,
               aes(x = x_left,
                   xend = x_left,
                   y = 2 - bar_height,
                   yend = 1 + bar_height),
               color = "gray70",
               linewidth = 0.4) +
  geom_segment(data = data.zoom,
               aes(x = x_right,
                   xend = 1,
                   y = 2 - bar_height,
                   yend = 1 + bar_height),
               color = "gray70",
               linewidth = 0.4) +
  # top bar labels
  geom_text(data = top_bar_labels,
            aes(x = x,
                y = 2 + bar_height + 0.08,
                label = label,
                hjust = hjust,
                color = label_color),
            vjust = 0,
            size = 3,
            fontface = "bold") +
  # bottom bar labels
  geom_text(data = bottom_bar_labels,
            aes(x = xlabel,
                y = 0.45,
                label = toupper(feature_type),
                color = label_color),
            fontface = "bold",
            size = 3) +
  # formatting
  facet_wrap(~ condition, nrow = 1,
             labeller = as_labeller(condition_labels)) +
  scale_fill_manual(values = all_colors) +
  scale_color_identity() +
  scale_y_continuous(limits = c(0.3, 2.7)) +
  coord_cartesian(clip = "off") +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text = element_markdown(
          size = 12,
          margin = margin(t = 0, b = 0.1, unit = "cm")),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        plot.margin = margin(0, 0, 0, 0),
        panel.spacing = unit(1, "lines"))
```

#### Endorse density plot
```{r}
data.endorse.summary = as.data.frame(
  emmeans(lmer(response ~ feature_type + (1 | id),
               data = data.endorse),
          ~ feature_type)) %>%
  rename(mean = emmean,
         ci_low = asymp.LCL,
         ci_high = asymp.UCL) %>%
  select(feature_type, mean, ci_low, ci_high)

p_endorse = ggplot(data.endorse,
                   aes(x = feature_type,
                       y = response)) +
  geom_violin(aes(fill = feature_type),
              color = NA,
              alpha = 0.3) +
  geom_pointrange(data = data.endorse.summary,
                  aes(y = mean,
                      ymin = ci_low,
                      ymax = ci_high,
                      color = feature_type),
                  size = 0.2,
                  linewidth = 0.7) +
  facet_wrap(~ "endorse",
             labeller = as_labeller(condition_labels)) +
  scale_fill_manual(values = feature_type_colors) +
  scale_color_manual(values = feature_type_colors) +
  scale_x_discrete(labels = c(
    "generic" = "generic",
    "demographics" = "demogr.",
    "dispositions" = "dispos.")) +
  scale_y_continuous(limits = c(0, 110),
                     breaks = c(0, 50, 100)) +
  coord_cartesian(ylim = c(0, 110)) +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text = element_markdown(
          size = 12,
          margin = margin(t = 0, b = 0.1, unit = "cm")),
        axis.title = element_blank(),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        plot.margin = margin(0, 0, 0, 0.5, unit = "lines"))
```

#### Coefficient plots

Odds ratios (for undo, same, and different):
```{r}
p_odds = ggplot(CI_95_all %>%
                  filter(condition != "endorse"),
                aes(x = predictor,
                    y = mean)) +
  geom_hline(yintercept = 1,
             linetype = "dashed",
             color = "grey50") +
  geom_pointrange(aes(ymin = l95, ymax = u95),
                  size = 0.3,
                  linewidth = 0.7,
                  color = 'gray30') +
  coord_flip(clip = "off") +
  facet_wrap(~ condition, nrow = 1,
             labeller = as_labeller(condition_labels)) +
  scale_y_continuous(name = "odds ratio",
                     limits = c(-0.5, 5.25),
                     breaks = seq(0, 5)) +
  theme(element_text(size = 12),
        legend.position = "none",
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text = element_text(
          size = 12, color = "black"),
        panel.border = element_rect(
          color = "black", fill = NA, linewidth = 0.4),
        panel.grid.major.x = element_line(),
        panel.grid.major.y = element_line(),
        panel.spacing = unit(1, "lines"),
        plot.margin = margin(0.3, 0, 0, 0, unit = "cm"))
```

Coefficients (for endorse):
```{r}
p_coeff = ggplot(CI_95_all %>%
                   filter(condition == "endorse"),
                 aes(x = predictor,
                     y = mean)) +
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "grey50") +
  geom_pointrange(aes(ymin = l95, ymax = u95),
                  size = 0.3,
                  linewidth = 0.7,
                  color = 'gray30') +
  coord_flip(clip = "off") +
  facet_wrap(~ condition, nrow = 1,
             labeller = as_labeller(condition_labels)) +
  scale_y_continuous(name = "coefficient",
                     limits = c(-2, 12.5),
                     breaks = c(0, 3, 6, 9, 12)) +
  theme(element_text(size = 12),
        legend.position = "none",
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(
          size = 12, color = "black"),
        panel.border = element_rect(
          color = "black", fill = NA, linewidth = 0.4),
        panel.grid.major.x = element_line(),
        panel.grid.major.y = element_line(),
        panel.spacing = unit(1, "lines"),
        plot.margin = margin(0.3, 0, 0, 1, unit = "lines"))
```

#### Model scatterplots

Undo, same, and different conditions:
```{r}
p_scatter_left = ggplot(model_predictions %>%
                          filter(condition != "endorse"),
                        aes(x = predicted,
                            y = actual)) +
  geom_abline(slope = 1,
              intercept = 0,
              linetype = "dashed",
              color = "grey50") +
  geom_point(alpha = 0.5, size = 1.5) +
  geom_text(data = model_metrics %>%
              filter(condition != "endorse"),
            aes(x = 0, y = 1.05, label = label),
            hjust = 0, vjust = 1,
            size = 3.5, lineheight = 0.9) +
  facet_wrap2(~ condition, nrow = 1,
              axes = "y", remove_labels = "x") +
  scale_x_continuous(name = "predicted selection",
                     breaks = c(0, 0.5, 1)) +
  scale_y_continuous(name = "selection",
                     breaks = c(0, 0.5, 1)) +
  coord_fixed(ratio = 1,
              xlim = c(0, 1),
              ylim = c(0, 1),
              clip = "off") +
  theme(text = element_text(size = 12),
        strip.background = element_blank(),
        strip.text = element_blank(),
        plot.margin = margin(0.3, 0, 0, 0, unit = "cm"),
        panel.spacing = unit(1, "lines"))
```

Endorse condition:
```{r}
p_scatter_right = ggplot(model_predictions %>%
                           filter(condition == "endorse"),
                         aes(x = predicted,
                             y = actual)) +
  geom_abline(slope = 1,
              intercept = 0,
              linetype = "dashed",
              color = "grey50") +
  geom_point(alpha = 0.5, size = 1.5) +
  geom_text(data = model_metrics %>%
              filter(condition == "endorse"),
            aes(x = 0, y = 105, label = label),
            hjust = 0, vjust = 1,
            size = 3.5, lineheight = 0.9) +
  facet_wrap(~ condition, nrow = 1) +
  scale_x_continuous(name = "predicted agreement",
                     breaks = c(0, 50, 100)) +
  scale_y_continuous(name = "agreement",
                     breaks = c(0, 50, 100)) +
  coord_fixed(ratio = 1,
              xlim = c(0, 100),
              ylim = c(0, 100),
              clip = "off") +
  theme(text = element_text(size = 12),
        strip.background = element_blank(),
        strip.text = element_blank(),
        panel.spacing = unit(1, "lines"),
        plot.margin = margin(0.3, 0, 0, 1, unit = "lines"))
```

#### Combining plots
```{r fig.height = 6, fig.width = 11}
(p_selections + (p_endorse / plot_spacer() + plot_layout(heights = c(20, 1))) +
  plot_layout(nrow = 1, widths = c(3.5, 1))) /
  (p_odds + p_coeff +
    plot_layout(nrow = 1, widths = c(3, 1))) /
  (p_scatter_left + p_scatter_right +
    plot_layout(nrow = 1, widths = c(3.5, 1))) +
  plot_layout(heights = c(1.25, 1, 2.5))

# ggsave(str_c(figures_dir, "exp2.pdf"),
#        width = 11, height = 6)
```

# Session info
```{r}
sessionInfo()
```

